### 2020-10-29 docs/ms/ms.RMD
Initial commit

---
title: "Manuscript title here"
author: "James R Junker"
date: Dept. of Ecology, Montana State University, Bozeman, MT USA
output:
   word_document:
     reference_docx: docx_template.docx
   pdf_document:
     keep_tex: true
   html_document: default
header-includes:
- \usepackage{lineno}
- \linenumbers
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

<!--- this is a note in Rmarkdown --->

Abstract
================================================================================

Introduction
================================================================================

Methods
================================================================================

Results
================================================================================

Discussion
================================================================================

References
================================================================================

<div id="refs"></div>

\clearpage