### 2021-07-01 doc/ms/tables-figures.Rmd

---
title: ""
author: "Report Author"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   word_document:
     reference_docx: docx_template.docx
   pdf_document:
     keep_tex: true
   html_document: default
header-includes:
- \usepackage{lineno}
- \usepackage{amsmath}
- \linenumbers
linestretch: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("./packages.R")
library(knitr)
```

## Tables
\newpage

Table 1.

```{r GINI table, echo = FALSE, warning=FALSE, message=FALSE}
drake::loadd(gini_analysis)
gini_analysis[["stream_gini_summary"]] %>% 
   dplyr::mutate(across(where(is.numeric),.fns = ~signif(.,2))) %>% 
   dplyr::mutate("Non-normalized Gini" = paste0(`Non-normalized Gini_mean`," ( ",`Non-normalized Gini_quant2.5`," - ", `Non-normalized Gini_quant97.5`," )"),
                 "Normalized Gini" = paste0(`Normalized Gini_mean`," ( ", `Normalized Gini_quant2.5`," - ", `Normalized Gini_quant97.5`," )")) %>% select(site, `Non-normalized Gini`,`Normalized Gini`) %>%
   dplyr::mutate(site = factor(site, levels = names(stream_order_list))) %>%
   knitr::kable(align = "lrr")

```

\newpage

Table 2.

\newpage

Figure descriptions

Figure 1. Species rank-flux plots among streams. Within each stream, species are ranked by descending annual total organic matter flux (mg AFDM $m^{-2} y^{-1}$).

```{r figure 1., echo = FALSE, warning = FALSE, message = FALSE, fig.width=5, fig.height=5, dpi = 450}
drake::loadd(annual_spp_flux_fig)
annual_spp_flux_fig
```

<!-- \newpage -->

<!-- Figure 2. Evenness profiles of species' relatives fluxes among streams. -->

<!-- ```{r evenness profile, echo=FALSE, warning=FALSE, message=FALSE,  fig.width=5, fig.height=5, dpi = 450} -->
<!-- loadd(hill_diversity_analysis) -->
<!-- hill_diversity_analysis -->

<!-- ``` -->

\newpage

Figure 3. Histograms of species trait distributions of (a) log~10~ annual production:biomass ratio (y^-1^), (b) log~10~ mean body mass (mg ind^-1^), and (c) log~10~ annual mean population biomass (mg m^-2^).

```{r trait hist, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, dpi = 450}
drake::loadd(spp_trait_histograms)
grid.draw(spp_trait_histograms)
```

\newpage

Figure 4. Relative cumulative fluxes among species in relation to ranked order of (a) annual production:biomass ratio (y^-1^) and (b) mean body size (mg ind^-1^).

```{r lorenz analysis, echo =FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, dpi = 450}
drake::loadd(lorenz_trait_fig)
grid.arrange(lorenz_trait_fig[["pb_lorenz_flux_plot"]], lorenz_trait_fig[["M_lorenz_flux_plot"]], ncol = 2, bottom = textGrob("Cumulative relative species", vjust = 1, gp = gpar(fontfamily = 'serif')),
                         left = textGrob("Cumulative relative flux", rot = 90, vjust = 1, gp = gpar(fontfamily = 'serif')))

```

\newpage

Figure 4. Bootstrapped distribution of skew statistics from OM fluxes with body size, *M*, and (b) population annual production:biomass ratio relative to the distribution of skew statistics of 100,000 random orderings.

```{r skew distribution, echo = FALSE, warning=FALSE, message=FALSE}
drake::loadd(random_skew_fig)
grid.arrange(random_skew_fig[[1]], random_skew_fig[[2]], ncol = 1)

```

\newpage

Figure 5. Non-random organization of community energy fluxes in relation to population biomass turnover (*P:B*) with mean annual temperature ($^\circ$C). Non-random organization is measured by the proportion of randomly generated skew statistics, $Sk_{flux}$, of equal or greater value. Observing an empirical value of $Sk_{flux}$ with a low probability of occurrence by random chance suggests non-random organization of community energy fluxes.

```{r echo = FALSE, warning=FALSE, message=FALSE}
drake::loadd(temperature_skew_fig)
temperature_skew_fig
```

\newpage
Figure S1. Modeled diet proportions among streams

```{r diets, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, dpi = 450}
drake::loadd(prelim_diet_figures)
prelim_diet_figures

```

\newpage

Figure S2. Lorenz plot of relative community flux by species in ascending order of annual population organic matter flux (mg AFDM m^-2^ y^-1^).

```{r lorenz flux, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, dpi = 450}
lorenz_trait_fig[["lorenz_flux_plot"]]

```

\newpage

Figure S3. Cumulative plot of relative community flux by species in relation to mean annual population biomass (mg m^-2^).

```{r biomass flux, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, dpi = 450}
lorenz_trait_fig[["bio_lorenz_flux_plot"]]
```